Volunteers in the Clouds: an Architecture for Low Cost and Potentially Massive Distributed Evolutionary Computation

### 2015-08-04 volunteer-clouds-ieee-ci-si.tex
Comentarios de la Introducción

\author{Juan-J.~Merelo*$^1$, Mario Garcia-Valdez$^2$, Pedro A. Castillo$^1$, Pablo Garc\'ia-S\'anchez$^1$
\thanks{Manuscript submitted for review on \today.}%
\thanks{$^1$Dept. of Computer Architecture and Technology and CITIC University of Granada}%
\thanks{$^2$Dept. of Graduate Studies at Instituto Tecnologico de Tijuana}%
\thanks{E-mail addresses: {\tt jmerelo@ugr.es}, {\tt
    mario@tectijuana.edu.mx}, {\tt pacv@ugr.es}, {\tt pablogarcia@ugr.es}}%
\thanks{*Corresponding author.}%
}

@@

There is, then, space for systems that do not have any of those
assumptions, such as the one presented in this paper, which will be
provisionally named {\sf NodIO}. NodIO is first an application that
runs in two tiers, a server and a client, the latter  can be possibly
run on a  % [Pedro] Both the server and the client in browsers?
          % different browsers as I understand? [JJ]  No, server in
          % server, client in browser, although there's nothing that
          % can't be done in a browser. 
browser. As such, it can be run in any computer that loads the web
page where the client is embedded, including possibly many over which
the scientist has no control.

Some time ago, our group faced  the problem of diminishing funds for
buying new     % [Pedro] This sentence is so so long. I would split it
               % in two sentences. [JJ] Done.
hardware and continuous hardware failures of existing
clusters. Considering this, we leveraged our experience of working
with web applications, including JavaScript, since their inception and other volunteer and
unconventional distributed evolutionary computing systems, and we 
decided to designd and relase a new free framework that would allow anyone to
create a volunteer distributed EC experiment using cloud resources as
servers and browsers as clients.

These systems must solve a series of issues before even starting to
measure performance; these issues are related to security as well as
safety, which introduces a {\em social} aspect to the design of a
computing system that makes it a {\em techno-social system}
\cite{vespignani2009predicting}. In this paper our intention is to
present the {\sf NodIO} framework and analyze it from two different
angles: the purely technical angle, including the decisions that went
into its features, and the social angle, to account on how the persons
that voluntarily choose to enter the web page that runs the system
determine its performance and how it can be modeled and, eventually,
predicted.

@@

% [Pedro] desde el abstract no se había nombrado "NodIO". 
% Creo que en este párrafo debería introducirse, principalmente el nombre.
% Añado una frase presentándolo:
In this sense, in this paper we propose {\sf NodIO}, a cloud or metal based 
volunteer computing system, whose architecture
has been developed using JavaScript on both the client and server sides.
All parts of the system are free and available with a free
license from \url{https://github.com/JJ/splash-volunteer}.
Thus, {\sf NodIO} has two parts:\begin{enumerate}
\item A REST server, that is, a server that includes several remote
  requests for storing and retrieving information (the CRUD cycle:
  create, request, update and delete) from the server. The information
  is of two kinds: {\em problem} based, that is, related to the
  evolutionary algorithm such as {\tt PUT}ing a chromosome in or {\tt
    GET}ing a random chromosome from it. It also peforms logging
  duties, but is basically a very lightweight and hign performance
  data store \cite{jj:idc:lowcost}. The server has the capability to
  run a single experiment, storing the chromosomes in a data structure
  that is reset when the solution is found. 
\item A client that includes the evolutionary algorithm as 
  JavaScript code embedded in a web page that displays graphs and some
  additional links and information on the experiment. This code runs
  an evolutionary algorithm {\em island} starting with random
  population and sending, every 100 generations, the best individual
  back to the server (via a {\tt PUT} request) and requesting a random
  individual from the server (via a {\tt GET} request). Apart from
  these constants, the algorithm can be run in many different ways:
  stopping when solution is found or, as we have done in this paper,
  using WebWorkers, which run in the background even if the webpage or
  browser tab with the experiment is not currently on focus.  
\end{enumerate}

@@

\begin{figure}[!t]
\centering
\includegraphics[width=12cm]{img/baseline-times.png}
\caption{Comparison of time to solution for the baseline system, using a {\tt
    node.js} client, for different population sizes. They only compaonre
runs in which solution has been found.}
\label{fig:baseline}
\end{figure}
First we establish the baseline performance by running the
evolutionary algorithm in a desktop client written using {\sf NodEO},
the node.js evolutionary algorithm library. This experiment tries to
find the solution to the 40-trap function with parameters $l=4, a=1,
b=2, z=3$, a population size equal 512, which was found in previous
experiments to be the correct one for finding the solution 95\% of the
time. The algorithm was run until the solution (a string with all
ones) was found or five millions evaluations had been performed. It
took around a minute, on average, that is $68.9694$ seconds, to
perform the fifty runs. In this experiment only 33, that is, 66\% were
successful. The experiments were repeated for population $p=1024$,
with success rate upgraded to 100\% and average duration $3.46
seconds$. Results for these two experiments 
are shown in Figure \ref{fig:baseline}.

>>>>

\author{Juan-J.~Merelo*$^1$, Mario Garc\'ia-Valdez$^2$, Pedro A. Castillo$^1$, Pablo Garc\'ia-S\'anchez$^1$
\thanks{Manuscript submitted for review on \today.}%
\thanks{$^1$Dept. of Computer Architecture and Technology and CITIC University of Granada}%
\thanks{$^2$Dept. of Graduate Studies at Instituto Tecnol\'ogico de Tijuana}%
\thanks{E-mail addresses: {\tt jmerelo@ugr.es}, {\tt
    mario@tectijuana.edu.mx}, {\tt pacv@ugr.es}, {\tt pablogarcia@ugr.es}}%
\thanks{*Corresponding author.}%
}

@@

There is, then, space for systems that do not have any of those %[Mario] these assumptions? 
assumptions, such as the one presented in this paper, which will be
provisionally named {\sf NodIO}. NodIO is first an application that
runs in two tiers, a server and a client, the latter  can be possibly
run on a  % [Pedro] Both the server and the client in browsers?
          % different browsers as I understand? [JJ]  No, server in
          % server, client in browser, although there's nothing that
          % can't be done in a browser. 
browser. As such, it can be run in any computer that loads the web
page where the client is embedded, including possibly many over which%[Mario]Quitar el possibly?   
the scientist has no control.

Some time ago, our group faced the problem of diminishing funds for
buying new     % [Pedro] This sentence is so so long. I would split it
               % in two sentences. [JJ] Done.
hardware and continuous hardware failures of existing
clusters. 
%[Mario] Otra opción: Some time ago, our group faced the problem of diminishing funds 
%for buying new hardware. This was aggravated by the increasing maintance costs 
%resulting from the continuous failures of existing clusters.
Considering this, we leveraged our experience of working
with web applications, including JavaScript, since their inception and other volunteer and
unconventional distributed evolutionary computing systems, and we 
decided to designd and relase a new free framework that would allow anyone to
create a volunteer distributed EC experiment using cloud resources as
servers and browsers as clients.

These systems must solve a series of issues before even starting to
measure their performance; these issues are related to security as well as
safety, which introduces a {\em social} aspect to the design of a
computing system that makes it a {\em techno-social system}
\cite{vespignani2009predicting}. In this paper our intention is to
present the {\sf NodIO} framework and analyze it from two different
angles: the purely technical angle, including the decisions that went
into its features, and the social angle, to account on how the persons
that voluntarily choose to enter the web page that runs the system
determine its performance and how it can be modeled and, eventually,
predicted.

@@

% [Pedro] desde el abstract no se había nombrado "NodIO". 
% Creo que en este párrafo debería introducirse, principalmente el nombre.
% Añado una frase presentándolo:
% [Mario] Nota: se menciona también en el segundo párrafo de la Introducción.
In this sense, in this paper we propose {\sf NodIO}, a cloud or metal based 
volunteer computing system, whose architecture
has been developed using JavaScript on both the client and server sides.
All parts of the system are free and available with a free
license from \url{https://github.com/JJ/splash-volunteer}.
Thus, {\sf NodIO} has two parts:\begin{enumerate}
\item A REST server, that is, a server that includes several remote %remote requests suena raro, propongo algo como: remote procedures o web interfaces   
  requests for storing and retrieving information (the CRUD cycle:
  create, request, update and delete) from the server. The information
  is of two kinds: {\em problem} based, that is, related to the
  evolutionary algorithm such as {\tt PUT}ing a chromosome in or {\tt
    GET}ing a random chromosome from it. It also peforms logging
  duties, but is basically a very lightweight and hign performance
  data store \cite{jj:idc:lowcost}. 
  %The information is of two kinds: Problem Based y ¿Cual otra? ¿Logging? No se entiende bien. ¿Podría ser problem domain?   
  The server has the capability to
  run a single experiment, storing the chromosomes in a data structure
  that is reset when the solution is found. 
\item A client that includes the evolutionary algorithm as 
  JavaScript code embedded in a web page that displays graphs and some
  additional links and information on the experiment. This code runs
  an evolutionary algorithm {\em island} starting with random
  population and sending, every 100 generations, the best individual
  back to the server (via a {\tt PUT} request) and requesting a random
  individual from the server (via a {\tt GET} request). Apart from
  these constants, the algorithm can be run in many different ways:
  stopping when solution is found or, as we have done in this paper,
  using WebWorkers, which run in the background even if the webpage or
  browser tab with the experiment is not currently on focus.  
\end{enumerate}

@@

\begin{figure}[!t]
\centering
\includegraphics[width=12cm]{img/baseline-times.png}
\caption{Comparison of time to solution for the baseline system, using a {\tt
    node.js} client, for different population sizes. They only compare on
runs in which solution has been found.}
\label{fig:baseline}
\end{figure}
First we establish the baseline performance by running the
evolutionary algorithm in a desktop client written using {\sf NodEO},
the node.js evolutionary algorithm library. This experiment tries to
find the solution to the 40-trap function with parameters $l=4, a=1,
b=2, z=3$, a population size equal 512, which was found in previous
experiments to be the correct one for finding the solution 95\% of the
time. The algorithm was run until the solution (a string with all
ones) was found or five millions evaluations had been performed. It
took around a minute, on average, that is $68.9694$ seconds, to
perform the fifty runs. In this experiment only 33, that is, 66\% were
successful. The experiments were repeated for population $p=1024$,
with success rate upgraded to 100\% and average duration $3.46
seconds$. Results for these two experiments 
are shown in Figure \ref{fig:baseline}.
